Coding Challenge #1: OLS
========================================================

The bulk of this code is from [Stpehanie Renfro's](mailtto:renfro@ohsu.edu) file `Challenge1R.R`.

Details
-------
> From: John McConnell  
> Sent: Tuesday, December 03, 2013 9:18 AM  
> To: Aaron Mendelson; Benjamin Chan; Hyunjee Kim; Jonah Todd-Geddes; Lorie Jacob; Margaret Sarna-Wojcicki; Peter Graven; Ruth Rowland; Stephanie Renfro  
> Subject: Developing programming examples for R, Stata, SAS
> 
> Group
>                                                                                                        
> Following up on our discussion last week, I propose the following exercise. I have tried to keep this very simple. We can get more involved if we think that would help. This is a baby step.
> 
> Let's have these people take the lead:
> 
> Stata -- Peter  
> R -- Stephanie  
> SAS -- Ben  
> 
> Let's work off this directory  
> X:\OHSU Shared\Restricted\OCHSER\PROJECTS\Code_Sharing  
> Using the file "Inpatient_Claims_Sample_1." (I have SAS, Stata, and R versions there).
> 
> Please create a word document and a statistical code file (e.g., a Stata "do" file or analog in SAS or R). The word document should describe, briefly
> a.  How you start the program  
> b.	How you locate the file Inpatient_Claims_Sample_1 within your program (or perhaps reference this in the code file).  
> c.	How you run your code  
> d.	How you view your results  
> 
> It is ok if the word file is only 4 sentences! I prefer to err on the side of keeping it simple at this point. We can always go back and add more detail later. This is not a manual.
> 
> The code file should be commented and describe how to do the following  
> a.	Load the dataset  
> b.	Describe the variables that exist in the data set (i.e., print out the names of the variables)  
> c.	Take the mean of CLM_PMT_AMT  
> d.	Run an OLS regression of CM_PMT_AMT  on CLM_UTLZTN_DAY_CNT and CLM_PASS_THRU_PER_DIEM_AMT (coefficient on  CLM_UTLZTN_DAY_CNT  should be about 489 and R-squared should be about 0.11)  
> e.	Save the output (the coefficient estimates, standard errors, p-values) in a text file or other file that can be opened up easily in Word or Notepad or a browser.  
> 
> When you're ready, send your Word doc and Code file to the group
> 
> WHEN?
> 
> By January 3, 2014
> 
> Ben, Peter, Stephanie: let me know by the end of the week if you'd like to change this plan (I.e., if it seems confusing, useless, or you're too busy).
> 
> Also, Ben made the following observations, which I agree with -- but I thought they might add to the complexity and so I thought we could revisit these at some point in the future, and for now just keep it very very simple.
> 
> COMMENTS FROM BEN
> 
> I think it might actually be instructive to see how to get the data from CSV to a native format for analysis (how easy/hard it is, how much can be automated, how much needs to be programmed). That might be a big enough task just to make it the first "project".
> 
> We'll want to put data on the X-drive. Probably a subfolder in the PROJECTS folder called "CodeSharing" or something like that and further organized with the following folders  
> 1.	Data  
> 2.	Task01ImportingData  
> 3.	Task02ComputingUnivariateStats  
> 4.	Task03ModelingOLS  
> 5.	Etc  


A. Load the dataset
-------------------
First set the working directory. The data file is 2 folders up from this Markdown file.
```{r}
# setwd("X:/OHSU Shared/Restricted/OCHSER/PROJECTS/Code_Sharing")
pathData <- file.path("..", "..")
```
Then read in the .csv file.
```{r}
f <- file.path(pathData, "DE1_0_2008_to_2010_Inpatient_Claims_Sample_1.csv")
sample <- read.csv(f)
```
Get the number of rows and columns in the data.
```{r}
dim(sample)
```


B. Print the names of the variables in the data
-----------------------------------------------
```{r}
names(sample)
```
Also display the first few rows
```{r}
head(sample)
```


C. Take the mean of `CLM_PMT_AMT`
---------------------------------
There are several ways to do this in R...
* Reference the desired data frame column by where it appears in the data. We know `CLM_PMT_AMT` is the 7th column based on the `names` output above.
```{r}
mean(sample[[7]]) 		
```
* Reference the desired data frame column by name. Note R variable names are **case-sensitive** (as are functions/commands/etc).
```{r}
mean(sample[["CLM_PMT_AMT"]])
``` 
* Similar to 2. but uses the single square bracket operator. The empty space before the comma signifies a wildcard match for the row position; i.e., return **all** the rows in `sample` but only the `CLM_PMT_AMT` column, then take the mean.
```{r}
mean(sample[, "CLM_PMT_AMT"])	 
```
* Use the dollar sign ($) to retrieve the desired data frame column. Note quotation marks aren't needed with this syntax.
```{r}
mean(sample$CLM_PMT_AMT) 	
```


D. Run an OLS regression
------------------------
Model `CLM_PMT_AMT` on `CLM_UTLZTN_DAY_CNT` and `CLM_PASS_THRU_PER_DIEM_AMT`.

Assign the name `results` to the outcome.
```{r}
results <- lm(CLM_PMT_AMT ~ CLM_UTLZTN_DAY_CNT + CLM_PASS_THRU_PER_DIEM_AMT, data=sample)
```
Print a summary of the results.
```{r}
summary(results)
```


E. Save the output
------------------
Save the output (coefficient estimates, standard errors, and p-values) in a text file.

Convert the output to a text string with the assigned name `out`.
```{r}
out <- capture.output(summary(results))
```
Write the output to a text file called `Output_R.txt`. The file will be written to the working directory you set earlier. Note, `write.csv` is similar to the `read.csv` function used earlier.
```{r}
write.csv(out, file="Output_R.txt")
```
